### 2016-07-27 report/Report/para/3.2-deeplearning4j.tex

\subsection{Deeplearning4j}
Deeplearning4j is a Java-based deep learning library built and supported by Skymind, a machine learning intelligence company, in 2014. It is an open source product designed for adoptability in industry, where Java is very common. The framework currently interfaces with both Java and Scala, with a Python SDK in-progress. Programming is primarily declarative, involving specifying network hyperparameters and layer information. Deeplearning4j integrates with Hadoop and Spark, or Akka and AWS for processing backends. Distributed execution provides data parallelism through the Iterative MapReduce model \cite{Itera38:online}. Each worker processes its own minibatch of training data, with workers periodically "reducing" (averaging) their parameter data. Deeplearning4j hosts its own Java linear algebra library, which claims 2x or more speedup over Python's Numpy library on large matrix multiplies \cite{ND4J'34:online}. Fault tolerance is not mentioned, although Spark does have built-in fault tolerance mechanisms.