### 2016-07-28 report/Report/para/0-abstract.tex
finished revisions

\begin{abstract}
The study and adoption of deep learning methods has led to significant progress in different application domains. As deep learning continues to show promise and its utilization matures, so does the infrastructure and software needed to support it. Various frameworks have been developed in recent years to facilitate both implementation and training of deep learning networks. As deep learning has also evolved to becoming distributed, there's a growing need for frameworks that can support execution across multiple machines. While deep learning frameworks restricted to running on a single machine have been studied and compared, frameworks which support distributed deep learning are relatively less known and well-studied. This paper seeks to bridge that gap by surveying, summarizing, and comparing frameworks which currently support distributed execution, including but not limited to Tensorflow, CNTK, Deeplearning4j, MXNet, H2O, and CaffeOnSpark.
\end{abstract}